---
title: "Case Study 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r cars}
bank <- read.csv('bank-additional.csv', sep=";")
```

```{r}
library(ggplot2)
library(dplyr)
library(MASS); library(car); library(olsrr)
library(DescTools);library(ResourceSelection)
library(caret);library(lattice);
library(gam);library(car)
library(ROCR);library(gridExtra)
library(gmodels)
```

# overview:

```{r pressure, echo=FALSE}
str(bank)
head(bank)
```

```{r}
summary(bank)
```
```{r}
#bank$job<-as.numeric(as.factor(bank$job))
#bank$marital<-as.numeric(as.factor(bank$marital))
#bank$education<-as.numeric(as.factor(bank$education))
#bank$default<-as.numeric(as.factor(bank$default))
#bank$housing<-as.numeric(as.factor(bank$housing))
#bank$loan<-as.numeric(as.factor(bank$loan))
#bank$contact<-as.numeric(as.factor(bank$contact))
#bank$month<-as.numeric(as.factor(bank$month))
#bank$day_of_week<-as.numeric(as.factor(bank$day_of_week))
#bank$campaign<-as.numeric(as.factor(bank$campaign))
#bank$previous<-as.numeric(as.factor(bank$previous))
#bank$poutcome<-as.numeric(as.factor(bank$poutcome))
```


## Check missing values
```{r}
colSums(is.na(bank))
```

## check duplicate values
```{r}
sum(duplicated(bank))
```
## Target variable 

```{r}
table(bank$y)

```
Number of no's greater than yes so there is a data imbalance.

## Unknown values
In the dataset description we also saw that the data contains 'unknown' values, Lets take a look at that.
```{r}
table(bank== "unknown")
colSums(bank=="unknown")
```
So job, marital,education,default,housing and loan have unknown values. We can decide what to do with unknows after exploratory data analysis.

## Visualization and Analysis

## 1.Age

```{r}
summary(bank$age)
unique(bank$age)
cor(bank$age,as.numeric(as.factor(bank$y)))
```

The distribution of age shows that majority of the bank's customers are aged between 25 - 50. More specifically, 50 % are between 32-47. The median age is 38. So, the target customers of the bank might be middle-aged adults.

```{r}
p1= ggplot(bank) + geom_histogram(aes(x=age),color="black", fill="grey") +
  ylab('Count') +  xlab('Age') +  geom_vline(aes(xintercept = mean(age), color = "red")) +
  scale_x_continuous(breaks = seq(0,100,10)) +  theme_minimal()
p2 = ggplot(bank) + geom_boxplot(aes(x='', y=age))
p3=ggplot(data = bank, aes(x=age, fill=y)) + geom_bar() + guides()
grid.arrange(p1, p2,p3,ncol=2)
```

## 2.Job
```{r}
summary(bank$job)
unique(bank$job)
#table(bank$job)
cor(as.numeric(as.factor(bank$job)),as.numeric(as.factor(bank$y)))
```
```{r}
p1 <- ggplot(bank,aes(y))+geom_bar(aes(job))
p2 <- ggplot(data = bank, aes(x=job, fill=y)) + geom_bar() + guides()
grid.arrange(p1, p2, ncol = 2)
```
## job variable classification w.r.t target variable
```{r}
CrossTable(bank$job,bank$y,prop.t = FALSE,prop.c = FALSE)
```

## 3.Marital
```{r}
summary(bank$marital)
unique(bank$marital)
#table(bank$marital)
cor(as.numeric(as.factor(bank$marital)),as.numeric(as.factor(bank$y)))
```
```{r}
p1<-ggplot(bank,aes(y))+geom_bar(aes(marital))
p2<-ggplot(data = bank, aes(x=marital, fill=y)) + geom_bar() + guides()
grid.arrange(p1, p2, ncol = 2)
```
## marital variable classification w.r.t target variable
```{r}
CrossTable(bank$marital,bank$y,prop.t = FALSE,prop.c = FALSE)
```

## 4.Education

```{r}
summary(bank$education)
unique(bank$education)
cor(as.numeric(as.factor(bank$education)),as.numeric(as.factor(bank$y)))
```

```{r}
p1<-ggplot(bank,aes(y))+geom_bar(aes(education))
p2<-ggplot(data = bank, aes(x=education, fill=y)) + geom_bar() + guides()
grid.arrange(p1, p2, ncol = 2)
```
## education classification w.r.t target variable
```{r}
CrossTable(bank$education,bank$y,prop.t = FALSE,prop.c = FALSE)
```

## 5.default

```{r}
summary(bank$default)
unique(bank$default)
#table(bank$default)
cor(as.numeric(as.factor(bank$default)),as.numeric(as.factor(bank$y)))
```
```{r}
p1 <- ggplot(bank,aes(y))+geom_bar(aes(default))
p2 <- ggplot(data = bank, aes(x=default, fill=y)) + geom_bar() + guides()
grid.arrange(p1, p2, ncol = 2)
```
## default classification w.r.t target variable
```{r}
CrossTable(bank$default,bank$y,prop.t = FALSE,prop.c = FALSE)
```

## 6.Housing
```{r}
summary(bank$housing)
unique(bank$housing)
#table(bank$housing)
cor(as.numeric(as.factor(bank$housing)),as.numeric(as.factor(bank$y)))
```
```{r}
p1 <- ggplot(bank,aes(y))+geom_bar(aes(housing))
p2 <- ggplot(data = bank, aes(x=housing, fill=y)) + geom_bar() + guides()
grid.arrange(p1, p2, ncol = 2)
```
## housing classification w.r.t target variable
```{r}
CrossTable(bank$housing,bank$y,prop.t = FALSE,prop.c = FALSE)
```

## 7. Loan

```{r}
summary(bank$loan)
unique(bank$loan)
cor(as.numeric(as.factor(bank$loan)),as.numeric(as.factor(bank$y)))
```

```{r}
p1 <- ggplot(bank,aes(y))+geom_bar(aes(loan))
p2 <- ggplot(data = bank, aes(x=loan, fill=y)) + geom_bar()+ guides()
grid.arrange(p1, p2, ncol = 2)
```
## loan classification w.r.t target variable
```{r}
CrossTable(bank$loan,bank$y,prop.t = FALSE,prop.c = FALSE)
```

## 8.Contact

```{r}
summary(bank$contact)
unique(bank$contact)
cor(as.numeric(as.factor(bank$contact)),as.numeric(as.factor(bank$y)))
```

```{r}
p1 <- ggplot(bank,aes(y))+geom_bar(aes(contact))
p2 <- ggplot(data = bank, aes(x=contact, fill=y)) + geom_bar()+ guides()
grid.arrange(p1, p2, ncol = 2)
```
## contact classification w.r.t target variable
```{r}
CrossTable(bank$contact,bank$y,prop.t = FALSE,prop.c = FALSE)
```
## 9.Month
```{r}
summary(bank$month)
unique(bank$month)
cor(as.numeric(as.factor(bank$month)),as.numeric(as.factor(bank$y)))
```

```{r}
p1 <- ggplot(bank,aes(y))+geom_bar(aes(month))
p2 <- ggplot(data = bank, aes(x=month, fill=y)) + geom_bar()+ guides()
grid.arrange(p1, p2, ncol = 2)
```
## month classification w.r.t target variable
```{r}
CrossTable(bank$month,bank$y,prop.t = FALSE,prop.c = FALSE)
```

## 10.Day_of_Week

```{r}
summary(bank$day_of_week)
unique(bank$day_of_week)
cor(as.numeric(as.factor(bank$day_of_week)),as.numeric(as.factor(bank$y)))
```

```{r}
p1 <- ggplot(bank,aes(y))+geom_bar(aes(day_of_week))
p2 <- ggplot(data = bank, aes(x=day_of_week, fill=y)) + geom_bar()+ guides()
grid.arrange(p1, p2, ncol = 2)
```
## days of week classification w.r.t target variable
```{r}
CrossTable(bank$day_of_week,bank$y,prop.t = FALSE,prop.c = FALSE)
```

## 11.Duation

```{r}
summary(bank$duration)
unique(bank$duration)
cor(as.numeric(as.factor(bank$duration)),as.numeric(as.factor(bank$y)))
```

```{r}
p1= ggplot(bank) + geom_histogram(aes(x=duration),color="black", fill="grey",bins=30) +
  ylab('Count') +  xlab('Duration') +  geom_vline(aes(xintercept = mean(duration), color = "red")) +
  scale_x_continuous(breaks = seq(0,100,10)) +  theme_minimal()
p2 = ggplot(bank) + geom_boxplot(aes(x='', y=duration))
p3=ggplot(data = bank, aes(x=duration, fill=y)) + geom_bar() + guides()
grid.arrange(p1, p2,p3,ncol=2)
```


## 12.Campaign

```{r}
summary(bank$campaign)
unique(bank$campaign)
cor(as.numeric(as.factor(bank$campaign)),as.numeric(as.factor(bank$y)))
```


```{r}
p1 <-  ggplot(bank) + geom_histogram(aes(x=campaign),color="black", fill="grey",bins=30) +
    ylab('Count') +  xlab('Campaign') +  geom_vline(aes(xintercept = mean(campaign), color = "red")) +
    scale_x_continuous(breaks = seq(0,100,10)) +  theme_minimal()
p2 <- ggplot(bank) + geom_boxplot(aes(x='', y=campaign))
p3 <- ggplot(data = bank, aes(x=campaign, fill=y)) + geom_bar() + guides()
grid.arrange(p1, p2,p3,ncol=2)
```

## 13.Pdays

```{r}
summary(bank$pdays)
unique(bank$pdays)
cor(as.numeric(as.factor(bank$pdays)),as.numeric(as.factor(bank$y)))
```

```{r}
bank <- bank %>% mutate(pdays = if_else(pdays == 999, "0", "1"))
#hist(as.numeric(bank$pdays))
p1 <- ggplot(data = bank, aes(x=pdays, fill=y)) + geom_bar()+ guides()
p1
```


## 14.Previous

```{r}
summary(bank$previous)
unique(bank$previous)
cor(as.numeric(as.factor(bank$previous)),as.numeric(as.factor(bank$y)))
```


```{r}
p1 <-  ggplot(bank) + geom_histogram(aes(x=previous),color="black", fill="grey",bins=30) +
    ylab('Count') +  xlab('Previous') +  geom_vline(aes(xintercept = mean(previous), color = "red")) +
    scale_x_continuous(breaks = seq(0,100,10)) +  theme_minimal()
p2 <- ggplot(data = bank, aes(x=previous, fill=y)) + geom_bar() + guides()
grid.arrange(p1,p2,ncol=2)
```
## 15.Poutcome

```{r}
summary(bank$poutcome)
unique(bank$poutcome)
cor(as.numeric(as.factor(bank$poutcome)),as.numeric(as.factor(bank$y)))
```

```{r}
p1 <- ggplot(bank,aes(y))+geom_bar(aes(poutcome))
p2 <- ggplot(data = bank, aes(x=poutcome, fill=y)) + geom_bar()+ guides()
grid.arrange(p1, p2, ncol = 2)
```

## poutcome classification w.r.t target variable
```{r}
CrossTable(bank$poutcome,bank$y,prop.t = FALSE,prop.c = FALSE)
```


## 16.emp.var.rate

```{r}
summary(bank$emp.var.rate)
unique(bank$emp.var.rate)
cor(as.numeric(as.factor(bank$emp.var.rate)),as.numeric(as.factor(bank$y)))
```

```{r}
p1 <-  ggplot(bank) + geom_histogram(aes(x=emp.var.rate),color="black", fill="grey",bins=30) +
    ylab('Count') +  xlab('emp.var.rate') +  geom_vline(aes(xintercept = mean(emp.var.rate), color = "red")) +
    scale_x_continuous(breaks = seq(0,100,10)) +  theme_minimal()
p2 <- ggplot(bank) + geom_boxplot(aes(x='', y=emp.var.rate))
p3 <- ggplot(data = bank, aes(x=emp.var.rate, fill=y)) + geom_bar() + guides()
grid.arrange(p1, p2,p3,ncol=2)
```
## 17.cons.price.idx

```{r}
summary(bank$cons.price.idx)
unique(bank$cons.price.idx)
cor(as.numeric(as.factor(bank$cons.price.idx)),as.numeric(as.factor(bank$y)))
```

```{r}
p1 <-  ggplot(bank) + geom_histogram(aes(x=cons.price.idx),color="black", fill="grey",bins=30) +
    ylab('Count') +  xlab('cons.price.idx') +  geom_vline(aes(xintercept = mean(emp.var.rate), color = "red")) +
    scale_x_continuous(breaks = seq(0,100,10)) +  theme_minimal()
p2 <- ggplot(bank) + geom_boxplot(aes(x='', y=cons.price.idx))
p3 <- ggplot(data = bank, aes(x=cons.price.idx, fill=y)) + geom_bar() + guides()
grid.arrange(p1, p2,p3,ncol=2)
```
## 18.cons.conf.idx

```{r}
summary(bank$cons.conf.idx)
unique(bank$cons.conf.idx)
cor(as.numeric(as.factor(bank$cons.conf.idx)),as.numeric(as.factor(bank$y)))
```

```{r}
p1 <-  ggplot(bank) + geom_histogram(aes(x=cons.conf.idx),color="black", fill="grey",bins=30) +
    ylab('Count') +  xlab('cons.conf.idx') +  geom_vline(aes(xintercept = mean(cons.conf.idx), color = "red")) +
    scale_x_continuous(breaks = seq(0,100,10)) +  theme_minimal()
p2 <- ggplot(bank) + geom_boxplot(aes(x='', y=cons.conf.idx))
p3 <- ggplot(data = bank, aes(x=cons.conf.idx, fill=y)) + geom_bar() + guides()
grid.arrange(p1, p2,p3,ncol=2)
```
## 19.euribor3m

```{r}
summary(bank$euribor3m)
unique(bank$euribor3m)
cor(as.numeric(as.factor(bank$euribor3m)),as.numeric(as.factor(bank$y)))
```

```{r}
p1 <-  ggplot(bank) + geom_histogram(aes(x=euribor3m),color="black", fill="grey",bins=30) +
    ylab('Count') +  xlab('euribor3m') +  geom_vline(aes(xintercept = mean(euribor3m), color = "red")) +
    scale_x_continuous(breaks = seq(0,100,10)) 
p2 <- ggplot(bank) + geom_boxplot(aes(x='', y=euribor3m))
p3 <- ggplot(data = bank, aes(x=euribor3m, fill=y)) + geom_bar(width=0.1) + guides()
grid.arrange(p1, p2,p3,ncol=2)
```

## 20.nr.employee

```{r}
summary(bank$nr.employed)
unique(bank$nr.employed)
cor(as.numeric(as.factor(bank$nr.employed)),as.numeric(as.factor(bank$y)))
```

```{r}
p1 <-  ggplot(bank) + geom_histogram(aes(x=nr.employed),color="black", fill="grey",bins=30) +
    ylab('Count') +  xlab('nr.employed') +  geom_vline(aes(xintercept = mean(nr.employed), color = "red")) +
    scale_x_continuous(breaks = seq(0,100,10)) 
p2 <- ggplot(bank) + geom_boxplot(aes(x='', y=nr.employed))
p3 <- ggplot(data = bank, aes(x=nr.employed, fill=y)) + geom_bar() + guides()
grid.arrange(p1, p2,p3,ncol=2)
```

## correlation and pairwise comparison 


```{r}
library(corrplot)
bank_mat <-bank
bank_mat$age<-as.numeric(as.factor(bank_mat$age))
bank_mat$job<-as.numeric(as.factor(bank_mat$job))
bank_mat$marital<-as.numeric(as.factor(bank_mat$marital))
bank_mat$education<-as.numeric(as.factor(bank_mat$education))
bank_mat$default<-as.numeric(as.factor(bank_mat$default))
bank_mat$duration<-as.numeric(as.factor(bank_mat$duration))
bank_mat$housing<-as.numeric(as.factor(bank_mat$housing))
bank_mat$loan<-as.numeric(as.factor(bank_mat$loan))
bank_mat$contact<-as.numeric(as.factor(bank_mat$contact))
bank_mat$pdays<-as.numeric(as.factor(bank_mat$pdays))
bank_mat$month<-as.numeric(as.factor(bank_mat$month))
bank_mat$day_of_week<-as.numeric(as.factor(bank_mat$day_of_week))
bank_mat$campaign<-as.numeric(as.factor(bank_mat$campaign))
bank_mat$previous<-as.numeric(as.factor(bank_mat$previous))
bank_mat$poutcome<-as.numeric(as.factor(bank_mat$poutcome))
bank_mat$y <- as.numeric(as.factor(bank_mat$y))
mat <- cor(bank_mat)

corrplot(mat,method="number",tl.cex=0.7,number.cex = 0.5,col=colorRampPalette(c("grey","blue","black"))(100))
```

```{r}
#bank$y <- ifelse(bank$y=='yes',1,0)
#bank$pdays<-as.numeric(as.factor(bank$pdays))
pairs(bank_mat)
```


```{r}
bank_1 <- bank_mat[, c('age','duration','month','campaign')]
pairs(bank_1)
```


## Model
```{r}
# Majority Null for pdays, duration should be removed
bank$pdays <- NULL 
bank$duration <- NULL

bank$y <- as.factor(bank$y)

# omitting null/unknowns
bank2 <- na.omit(bank)
bank2[bank2 == "unknown"] <- NA
bank2 <- na.omit(bank2)
str(bank2)


levels(as.factor(bank2$default))
bank2$default = as.factor(bank2$default)
summary(bank2)

# Splitting data train and test
splitBank = sort(sample(nrow(bank2), nrow(bank2)*.75))
Btrain <- bank2[splitBank,]
Btest <- bank2[-splitBank,]

# AIC model
form_2 = as.formula(paste0('y ~ .'))
form_2
set.seed(1234)
objControl <- trainControl(method = "none",
                           summaryFunction = twoClassSummary,
                           classProbs = TRUE,
                           savePredictions = TRUE)

Btrain[Btrain == "blue-collar"] <- "bluecollar"
Btrain[Btrain == "self-employed"] <- "selfemployed"

aicmodel <- train(form_2, data = Btrain,
               method = 'glmStepAIC',
               trControl = objControl,
               metric = "ROC",
               direction = 'forward')

summary(aicmodel)
g1 = glm(formula = y ~ ., data = Btrain, family = binomial)
summary(g1)
vif(g1)

#multicollinerty square last column > 10
g2 = glm(formula = y ~ age + job + marital + education + default + housing + loan + contact + month + day_of_week + campaign + previous + poutcome + cons.conf.idx, data = Btrain, family = binomial) 
summary(g2)
vif(g2)

# model 3 from AIC
g3 = glm(formula = y ~ nr.employed + poutcome + month + contact + cons.conf.idx + campaign, data = Btrain, family = binomial)
vif(g3)

# Probability for 1st model
Btrain$PredProb = predict.glm(g2, newdata = Btrain, type = "response")
Btrain$Predy = ifelse(Btrain$PredProb >= 0.5,"yes","no")
confusionMatrix(Btrain$y, as.factor(Btrain$Predy))

# ROC CURVE MODEL 1
lgPredObj <- prediction(Btrain$PredProb, Btrain$y)
lgPerfObj <- performance(lgPredObj, "tpr", "fpr")
plot(lgPerfObj, main = "ROC Curve", col = 2, lwd = 2)
abline(a = 0, b = 1, lwd = 2, lty = 3, col = "black")


find_cutoff <- function(actual_value, positive_class_name, negitive_class_name, pred_probability, pred_01=1, pred_10=1){
  # Initialising Variables
  msclaf_cost <- c()
  youden_index <- c()
  cutoff <- c()
  pred00 <- c() #correct classification of negative as negative (Sensitivity)
  pred01 <- c() #mis classification of negative to positive class (actual is 0, predicted 1)
  pred10 <- c() #mis classification of positive to negative class (actual 1 predicted 0)
  pred11 <- c() #correct classification of positive as positive (Specificity)
  
  costs = matrix(c(0, pred_01, pred_10, 0), ncol = 2)
  
  for (i in 1:100) {
    predList <- as.factor(ifelse(pred_probability >= i/100, positive_class_name, negitive_class_name))
    tbl <- table(predList, actual_value)
    
    # Classifying actual no as yes
    pred00[i] <- tbl[1]/(tbl[1] + tbl[2])
    
    pred01[i] <- tbl[2]/(tbl[1] + tbl[2])
    
    # Classifying actual yes as no
    pred10[i] <- tbl[3]/(tbl[3] + tbl[4])
    
    pred11[i] <- tbl[4]/(tbl[3] + tbl[4])
    
    cutoff[i] <- i/100
    msclaf_cost[i] <- pred10[i] * costs[3] + pred01[i] * costs[2]
    youden_index[i] <- pred11[i] + pred00[i] - 1
  }
  df.cost.table <- as.data.frame(cbind(cutoff, pred10, pred01, pred11, pred00, youden_index, msclaf_cost))
  cat(paste0('The ideal cutoff for:\n Yodens Index approach : ', which.max(df.cost.table$youden_index)/100))
  cat(paste0('\n Cost based approach : ', which.min(df.cost.table$msclaf_cost)/100))
  ggplot(df.cost.table, aes(x = cutoff)) +
    geom_line(aes(y = youden_index, color = 'yoden index')) +
    geom_line(aes(y = msclaf_cost, color = 'misclassification cost'))+
    labs(x = 'Cutoff p value', y='Index',  title = 'Cutoff p value',fill = 'Plot') +
    theme_minimal()+ theme(legend.position="bottom")
}

# CUTOFF MODEL
find_cutoff(actual_value = Btrain$y, positive_class_name = 'yes', negitive_class_name = 'no', pred_probability = Btrain$PredProb, pred_01 =3, pred_10 = 1)
Btrain$Predy = ifelse(Btrain$PredProb >= 0.14,"yes","no")
confusionMatrix(Btrain$y, as.factor(Btrain$Predy))

#### AIC MODEL
Btrain$PredProb3 = predict.glm(g3, newdata = Btrain, type = "response")
Btrain$Predy3 = ifelse(Btrain$PredProb3 >= 0.5,"yes","no")
confusionMatrix(Btrain$y, as.factor(Btrain$Predy3))


#Model 3 Cutoff
find_cutoff(actual_value = Btrain$y, positive_class_name = 'yes', negitive_class_name = 'no', pred_probability = Btrain$PredProb3, pred_01 =3, pred_10 = 1)
Btrain$Predy3 = ifelse(Btrain$PredProb3 >= 0.13,"yes","no")
confusionMatrix(Btrain$y, as.factor(Btrain$Predy3))
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
